---
title: "Titanic LM, CART, and RF"
navbar:
  title: "Simon Raymond"
  left:
    - text: "Home"
      href: index.html
    - text: "About"
      href: about.html
    - text: "Titanic XGBoosting"
      href: Titanic_XGBoosting.html
    - text: "Titanic Adaboost"
      href: Titanic_Adaboost.html
    - text: "Titanic Nnet"
      href: Titanic_Nnet.html
    - text: "Titanic_LM_CART_RF"
      href: Titanic_LM_CART_RF.html
      name: SimonRaymond 
---

titanic rf lm and cart

# Data Prep

Download the Data
```{r, warning=FALSE}
train_titanic <- read.csv("train.csv")
```

the columns i will use are y (Survived), Pclass, Sex, Age, SibSp, Parch, Fare, Emarked as they are the easiest to work with. other possibilities include using the titles for names 
```{r, warning=FALSE}
library(dplyr)
data <- train_titanic[ ,c(2, 3, 5, 6, 7, 8, 10, 12)]
data <- data %>% rename(y = Survived)
data$y <- factor(data$y)
```

Convert the characters to factors. 
```{r, warning=FALSE}
char_cols <- sapply(data, is.character)
data[, char_cols] <- lapply(data[, char_cols], factor)
```

We have NAs we need to deal with
```{r, warning=FALSE}
colSums(is.na(data))
```

Firsts i am going to predict the NAs for Age with a simple Random Forest model.
```{r, warning=FALSE}
na_index <- which(is.na(data$Age))#index the missing NAs
na_data <- data[na_index, ]#create a data frame of the NAs indexed
c_data <-data[-na_index, ]#take the complete data to c_data
na_data <- na_data[, -which(names(data) == "Age")] # take out the collum that only has NAs in it
```

 i will use na.roughfix() for the two missing Embarked values
```{r, warning=FALSE}
library(randomForest)
c_data <- na.roughfix(c_data)# there are two NAs in embarked
na_data <-  na.roughfix(na_data)
```

Predict our NAs
```{r, warning=FALSE}
library(randomForest)

model_rf <- randomForest(Age ~.,
                         ntree = 1200, 
                         data = c_data) 

pred_na <- predict(model_rf, na_data)
```

Bind our predictions to the corresponding index number for each missing NA in "data"
```{r, warning=FALSE}
na_n_ind <- cbind(pred_na, na_index)

data$Age[na_n_ind[, 2]] <- na_n_ind[, 1]
```

finnaly just use na.roughfix() for the two embarked values since it is so little
```{r, warning=FALSE}
library(randomForest)
data <- na.roughfix(data)
colSums(is.na(data))
```

Now do the same for our test set
```{r, warning=FALSE}
str(data)
```




# linear Model

Y has to be numeric for lm. also i am facing a issue where for some of the splits not all factor levels from embarked are being selected which will produce a error. since lm is more of a basline i will simply remove embarked. i know that glm exists but i prefer lm as a basline.
```{r, warning=FALSE}
data_lm <- data
data_lm$y <- as.numeric(data_lm$y)
data_lm$y <- data_lm$y - 1 # bring it back to 1 and 0s
data_lm <- data_lm[, -which(colnames(data_lm) == "Embarked")]

```

```{r, warning=FALSE}
library(foreach)
library(doParallel)

n <- 100 # this number times 3

obs <- nrow(data_lm) # observations

trials <- 1:3
numCores <- detectCores()-1

cl <- makeCluster(numCores)
registerDoParallel(cl)

lst <- foreach(k=trials, .packages = c("ROCR")) %dopar% {
  
  auc <- c()
  for (i in 1:n){
    
    idx <- sample(obs, obs, replace = TRUE)
    train <- data_lm[idx,]
    test <- data_lm[-idx, ]
    
    model <- lm(y ~ ., data = train)
    
    phat <- predict(model, test)

    pred_rocr <- prediction(phat, test$y)
    auc_ROCR <- performance(pred_rocr, measure = "auc")
    auc[i] <- auc_ROCR@y.values[[1]]
    
  }
  return(auc)
}

stopCluster(cl)

# combine the results
auc <- unlist(lst)

mean(auc)
sd(auc)

```

```{r, warning=FALSE}

# plot auc and mean
plot(auc, col="red")
abline(a = mean(auc), b = 0, col = "blue", lwd = 2)
abline(a = mean(auc)-1.96*sd(auc), b = 0, col = "green", lwd = 3)
abline(a = mean(auc)+1.96*sd(auc), b = 0, col = "green", lwd = 3)

```

# CART

This is using default params but once again it is mearly a baseline to examine
```{r, warning=FALSE}
library(rpart)
library(ROCR)
library(foreach)
library(doParallel)

n <-  100 # this number times 3
obs <- nrow(data) # observations

trials <- 1:3
numCores <- detectCores()-1

cl <- makeCluster(numCores)
registerDoParallel(cl)

lst <- foreach(k=trials, .packages = c("rpart", "ROCR")) %dopar% {
  
  auc <- c()
  for (i in 1:n){
    idx <- sample(obs, obs, replace = TRUE)
    train <- data[idx,]
    test <- data[-idx, ]
    
    model <- rpart(y ~ ., data = train, method = "class")
    
    phat <- predict(model, test, type = "prob")[,2]
    
    pred_rocr <- prediction(phat, test$y)
    auc_ROCR <- performance(pred_rocr, measure = "auc")
    auc[i] <- auc_ROCR@y.values[[1]]
    
  }
  return(auc)
}

stopCluster(cl)

# combine the results
auc <- unlist(lst)

mean(auc)
sd(auc)

```

```{r, warning=FALSE}

# plot auc and mean
plot(auc, col="red")
abline(a = mean(auc), b = 0, col = "blue", lwd = 2)
abline(a = mean(auc)-1.96*sd(auc), b = 0, col = "green", lwd = 3)
abline(a = mean(auc)+1.96*sd(auc), b = 0, col = "green", lwd = 3)

```


# Random Forest
This is a Random Forest Algorithm that uses Parallel processing. In the loop ntree is set to B. I use Bootstrapping data samples instead of k-fold-cross-validation as it is more robust and lets me reproduce the model as many times as i like.
```{r, warning=FALSE}



library(randomForest)
library(ROCR)
library(foreach)
library(doParallel)

B <- 1200
n <-  100 # this number times 3

obs <- nrow(data) # observations

trials <- 1:3
numCores <- detectCores()-1

cl <- makeCluster(numCores)
registerDoParallel(cl)


lst <- foreach(k=trials, .packages = c("randomForest", "ROCR")) %dopar% {
  
  auc <- c()
  for (i in 1:n){
    
    idx <- sample(obs, obs, replace = TRUE)
    train <- data[idx,]
    test <- data[-idx, ]
    
    model <- randomForest(y ~ ., ntree = B, data = train)
    
    phat <- predict(model, test, type = "prob")

  
    pred_rocr <- prediction(phat[,2], test$y)
    auc_ROCR <- performance(pred_rocr, measure = "auc")
    auc[i] <- auc_ROCR@y.values[[1]]
    
  }
  return(auc)
}

stopCluster(cl)

# combine the results
auc <- unlist(lst)

mean(auc)
sd(auc)
```

```{r, warning=FALSE}

# plot auc and mean
plot(auc, col="red")
abline(a = mean(auc), b = 0, col = "blue", lwd = 2)
abline(a = mean(auc)-1.96*sd(auc), b = 0, col = "green", lwd = 3)
abline(a = mean(auc)+1.96*sd(auc), b = 0, col = "green", lwd = 3)

```

